---
title: "STA325 Final Project-analysis"
author: "Alice Liao, Revo Tesha, Salvador, Cindy, Evelyn "
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(cowplot)
library(janitor)
library(nnet)
library(caret)
library(broom)
library(tree)
library(brms)
set.seed(10)
```

## Exploratory Data Analysis

### Training data
```{r}
#load the separated datset from other rmd
train = sample(max(dim(oh_wv_2012)), max(dim(oh_wv_2012))*0.75) # 75% training
train_oh_wv_all = oh_wv_2012[train,]
test_oh_wv_all = oh_wv_2012[-train,]
```

### Distribution of Variables (and possible transformations):

```{r}
# Dosage strength -- no transformation
ds_p = ggplot(oh_wv_all, aes(x = DOS_STR)) +
  geom_histogram() +
  labs(x = "Dosage Strength")
ds_p
```

```{r}
#CHANGE THIS-WE SHOULD USE FULL DATASET FOR EDA
# hydrocodone -- log transformation
hyd_wt_p = ggplot(train_oh_wv_all, aes(x = hyd_wt)) +
  geom_histogram() +
  labs(x = "Total Active Weight for Hydrocodone")

thyd_wt_p = ggplot(train_oh_wv_all, aes(x = log(hyd_wt))) +
  geom_histogram() +
  labs(x = "Log Total Active Weight for Hydrocodone")

plot_grid(hyd_wt_p, thyd_wt_p)
```

```{r}
# oxycodone -- log transformation
oxy_wt_p = ggplot(train_oh_wv_all, aes(x = oxy_wt)) +
  geom_histogram() +
  labs(x = "Total Active Weight for Oxycodone")

toxy_wt_p = ggplot(train_oh_wv_all, aes(x = log(oxy_wt))) +
  geom_histogram() +
  labs(x = "Log Total Active Weight for Oxycodone")

plot_grid(oxy_wt_p, toxy_wt_p)
```

```{r}
# number of distributors per 10,000 in a county -- log transformation
ndptt_p = ggplot(train_oh_wv_all, aes(x = number_of_distributors_ptt)) +
  geom_histogram() +
  labs(x = "Number of Distributors per 10,000 in a County")

tndptt_p = ggplot(train_oh_wv_all, aes(x = log(number_of_distributors_ptt))) +
  geom_histogram() +
  labs(x = "Log Number of Distributors per 10,000 in a County")

plot_grid(ndptt_p, tndptt_p)
```

```{r}
# number of pharmacies per 10,000 for a county -- no transformation necessary?
nptt_p = ggplot(train_oh_wv_all, aes(x = pharmacy_num_ptt)) +
  geom_histogram() +
  labs(x = "Number of Pharmacies per 10,000 in a County")

# tnptt_p = ggplot(train_oh_wv_all, aes(x = log(pharmacy_num_ptt))) +
#   geom_histogram() +
#   labs(x = "Log Number of Pharmacies per 10,000 in a County")

# plot_grid(nptt_p, tnptt_p)
```

```{r}
# pharma_num = ggplot(train_oh_wv_all, aes(x = pharmacy_num)) +
#   geom_histogram() +
#   labs(x = "Number of pharmacies")
# 
# log_pharma_num = ggplot(train_oh_wv_all, aes(x = log(pharmacy_num))) +
#   geom_histogram() +
#   labs(x = "Log Number of pharmacies")


# median household income -- log transformation necessary?
med_income = ggplot(train_oh_wv_all, aes(x = median_income)) +
  geom_histogram() +
  labs(x = "Median income")

log_med_income = ggplot(train_oh_wv_all, aes(x = log(median_income))) +
  geom_histogram() +
  labs(x = "Log median income")

plot_grid(med_income, log_med_income)
```

```{r}
# active drug weight per person in a county -- log transformation necessary?
actwt_pp_plot = ggplot(train_oh_wv_all, aes(x = act_wt_person_county)) +
  geom_histogram() +
  labs(x = "active weight per person")

log_actwt_pp_plot = ggplot(train_oh_wv_all, aes(x = log(act_wt_person_county))) +
  geom_histogram() +
  labs(x = "Log active weight per person")

plot_grid(actwt_pp_plot, log_actwt_pp_plot)
```

```{r}
# # distribution channel -- looks ok I guess.
# plot(train_oh_wv_all$most_dist_channel)
```

```{r}
# # market dominance -- also looks ok I guess.
# plot(train_oh_wv_all$dominance)
```

```{r}
# # market dominance -- also looks ok I guess.
# plot(train_oh_wv_all$political_aff)
```

### Predictors vs. Response:

Responses: Death Rate and Income:

I think we also need EDA for relationship between EACH predictor and response -- to help
justify how we treat each variable.

Response: Death Rate: (kinda messy since est_death_rate_cat is 'numerical' and categorical but also since nothing falls in between)
```{r}
oh_wv_2012 %>%
  ggplot(mapping = aes(x = dominance, y = est_death_rate_cat)) + geom_boxplot() + xlab("Dominance") + ylab("Estimated Death Rate Category") 


oh_wv_2012 %>%
  ggplot(mapping = aes(x = Population, y = est_death_rate_cat)) + geom_point()+ xlab("Population") + ylab("Estimated Death Rate Category") 


ggplot(mapping = aes(x = Political.affiliation, y = est_death_rate_cat), data = oh_wv_2012) + geom_boxplot() + xlab("Political Affiliation") + ylab("Estimated Death Rate Category")

oh_wv_2012 %>%
  ggplot(mapping = aes(x = dominance, y = est_death_rate_cat)) + geom_boxplot() + xlab("Dominance") + ylab("Estimated Death Rate Category") 

oh_wv_2012 %>%
  ggplot(mapping = aes(x = pharmacy_num, y = est_death_rate_cat)) + geom_point() + xlab("Number of Pharmacies") + ylab("Estimated Death Rate Category") 

oh_wv_2012 %>%
  ggplot(mapping = aes(x = all_active_wt, y = est_death_rate_cat)) + geom_point() + xlab("Active weight total") + ylab("Estimated Death Rate Category") 

oh_wv_2012 %>%
  ggplot(mapping = aes(x = most_dist_channel, y = est_death_rate_cat)) + geom_boxplot() + xlab("Most Distributed Channel") + ylab("Estimated Death Rate Category") 
  
```

Focus on Median Income here:

```{r}
oh_wv_2012 %>%
  ggplot(mapping = aes(x = Political.affiliation, y = median_income)) + geom_boxplot() + xlab("Political Affiliation") + ylab("Median Income")

oh_wv_2012 %>%
  ggplot(mapping = aes(x = dominance, y = median_income)) + geom_boxplot() + xlab("Dominance") + ylab("Median Income")


oh_wv_2012 %>%
  ggplot(mapping = aes(x = all_active_wt , y = median_income)) + geom_point()  + xlab("All Active Weight") + ylab("Median Income")

med_inc_mod3 <- lm(median_income ~ all_active_wt, data = oh_wv_2012)
plot(med_inc_mod3)


oh_wv_2012 %>%
  ggplot(mapping = aes(x = most_dist_channel , y = median_income)) +  geom_boxplot()  + xlab(" Most Distributed Channel") + ylab("Median Income")

oh_wv_2012 %>%
  ggplot(mapping = aes(x = pharmacy_num , y = median_income)) + geom_point()  + xlab("Number of Pharamacies") + ylab("Median Income")

med_inc_mod6 <- lm(median_income ~ pharmacy_num, data = oh_wv_2012)
plot(med_inc_mod6)


oh_wv_2012 %>%
  ggplot(mapping = aes(x = Population , y = median_income)) + geom_point()  + xlab("Population") + ylab("Median Income")

med_inc_mod7 <- lm(median_income ~ Population, data = oh_wv_2012)
plot(med_inc_mod7)

```

```{r}

```

```{r}

```

```{r}

```

Average active weight of drugs per person by county plotted on a map and compared to opioids deaths by county.
```{r} 
#change datset starting from here
ggplot(data = oh_wv_all) +
  geom_point(aes(x = act_wt_person_county, y = est_death_rate_cat)) +
  labs(x = "Total Active Weight of Drug Per Person 2006-2012",
       y = "Total Overdose Death Rate in a County 2006-2012")
```

Relationship between number of manufacturers/distributors shipping to a county vs opioid deaths by county.
```{r}
oh_wv_all %>%
  group_by(REPORTER_NAME, est_death_rate_cat, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY, est_death_rate_cat) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters)) %>%
  ggplot(., aes(x = est_death_rate_cat, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Average death rate of a county", y = "Number of manufacturers involved",
       title = "Relationship between number of manufacturers and death rates in a county?")
```

Number of manufacturers/distributors shipping to a county vs amount of opioids in active weight ordered by 
pharmacies in the county.
```{r}
oh_wv_all %>%
  group_by(REPORTER_NAME, TOTAL_ACTIVE_WT, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters),
            total_active_drug_for_given_county = sum(TOTAL_ACTIVE_WT)) %>%
  ggplot(., aes(x = total_active_drug_for_given_county, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Total Active Drug Weight", y = "Number of manufacturers shipping to the county",
       title = "Total active drug weight vs. number of manufacturers for that county")
```

Average active weight of drugs per person vs number of pharmacies per 10,000 people.
```{r}
pharmacies = oh_wv_all %>%
  distinct(BUYER_ADDRESS, .keep_all = TRUE) %>%
  group_by(BUYER_COUNTY) %>%
  summarize(number_of_pharmacies = n(), Population = first(Population))

active_weight = oh_wv_all %>%
  group_by(BUYER_COUNTY) %>%
  distinct(act_wt_person_county, .keep_all = TRUE) %>%
  ungroup() %>%
  dplyr::select(BUYER_COUNTY, act_wt_person_county)
  
pop_and_pharmacies = inner_join(active_weight, pharmacies) %>%
  ###### someone check if this is how you calculate number of pharmacies per 10,000 ######
  mutate(pharmacies_per_tt = (10000*number_of_pharmacies)/Population)

  ########################################################################################
ggplot(pop_and_pharmacies, aes(x = act_wt_person_county, y = pharmacies_per_tt)) +
geom_point() +
labs(x = "Avg. active weight of drugs per person", y = "Number of pharmacies")
```

<!-- Distribution of oxycodone and hydrocodone. -->
<!-- ```{r} -->
<!-- ``` -->

<!-- Average active weight of drugs per person vs strength of dose. -->
<!-- ```{r} -->
<!-- # someone has code for this I think -->
<!-- ``` -->

### Interactions:
```{r}
#CHANGE DATASET
I1 = ggplot(data=oh_wv_all, aes(x= median_income ,y=est_death_rate_cat,color=as.factor(Political.affiliation))) +
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Income vs Death Rate",
       x=" Median Income", y="Death Rate (Categorical)",
       color="Political Affiliation")

I2 = ggplot(data=train_oh_wv_all, aes(x= pharmacy_num ,y=median_income ,color= Political.affiliation)) +
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Median Income vs Number of Pharmacies", 
       x=" Number of Pharmacies", y="Median Income",
       color="Political Affiliation")

I3 <- ggplot(data=train_oh_wv_all, aes(x= pharmacy_num ,y=est_death_rate_cat ,color= Political.affiliation)) +
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Number of Pharmacies vs Death Rate", 
       x=" Number of Pharmacies", y="Estimated Death Rate",
       color="Political Affiliation")

I4 = ggplot(data=train_oh_wv_all, aes(x= act_wt_person_county ,y=est_death_rate_cat ,color= dominance)) +
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Active Drug Weight PP in a County vs Death Rate", 
       x="Active Drug Weight P", y="Estimated Death Rate",
       color="Market Dominance")

##########################################################################################################
#### Active weight person vs death rates by percentage of hydrocodone and oxycodone -- @Cindy? ############
##########################################################################################################

plot_grid(I1, I2, I3, I4)
```


<!-- ```{r} -->
<!-- # total active weight by county and drug name -->
<!-- arcos2 <- arcos1 %>% -->
<!--   group_by(BUYER_COUNTY,DRUG_NAME) %>% -->
<!--   summarise(all_active_wt = sum(TOTAL_ACTIVE_WT))  -->
<!-- head(arcos2) -->
<!-- ``` -->


## Modeling
```{r}
set.seed(12)
### Bayesian cumulative mixed-effects model:

# An example of what a mixed-effects model would look like (I don't think we need it,
# thought, especially if we don't use report name or manufacturer in our model or
# if we are interested in seeing the effects these two variables have on death
# rates). The reason we would want random effects is because the variables have
# too many levels to look at.

# Stuff looking like (1|...) are random effects. Everything else is 'fixed'.

# If you remove the random effect, this is just a standard multicategorical
# regression (bayesian approach though).

################################### change variables #############################################
# m.c_r.e_model = brm(est_death_rate_cat ~ QUANTITY + TOTAL_ACTIVE_WT + DOSAGE_UNIT +
#                       DOS_STR + EQU_MORPHINE + State + Population + median_income + 
#                       Political.affiliation + county_total_wt + act_wt_person_county + 
#                       (1|BUYER_COUNTY) + (1|REPORTER_NAME) + (1|MANUFACTURER), data = train_oh_wv_all,
#                     chains = 1, family = "cumulative", iter = 2, warmup = 0) # iter should be big.
# summary(m.c_r.e_model)
###################################################################################################
```

```{r}
### cumulative logistic regression
cumu.logistic = nnet::multinom(est_death_rate_cat ~ log(pharmacy_num) + most_dist_channel + 
                                 dominance + log(median_income) +  political_aff + log(act_wt_person_county), 
                               data=train_oh_wv_all)
summary(cumu.logistic)
mostImportantVariables <- varImp(cumu.logistic)
mostImportantVariables$Variables <- row.names(mostImportantVariables)
mostImportantVariables <- mostImportantVariables[order(-mostImportantVariables$Overall),]
print(head(mostImportantVariables))
knitr::kable(cumu.logistic %>% tidy(conf.int=TRUE),format="html",digits=3)
```


```{r}
#CHANGE DATASET
# Classification tree model
set.seed(13)
class.tree.1 <- tree(est_death_rate_cat ~ log(pharmacy_num) + most_dist_channel + dominance +
                       log(median_income) + political_aff + log(act_wt_person_county), data=train_oh_wv_all)
summary(class.tree.1)
plot(class.tree.1)
text(class.tree.1, pretty = 0)
class.tree.1 # Branches that lead to terminal nodes are indicated using asterisks
```