---
title: "STA325 Final Project"
author: "Alice Liao, Revo Tesha, Salvador, Cindy, Evelyn "
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(cowplot)
set.seed(10)
```

## Data

### Drug poisoning dataset
```{r}
# load original dataset
death.wv <- read.csv("data/Drug_Poisoning_Mortality_by_County.csv", header = TRUE) 

death.oh <- read.csv("data/NCHS_-_Drug_Poisoning_Mortality_by_County__United_Statesoh.csv", 
                     header = TRUE, as.is = TRUE) %>%
   mutate(Population = as.numeric(gsub(",","",Population,fixed=TRUE)))

death0<-rbind(death.wv, death.oh)%>%
    rename(est_death_rate = Estimated.Age.adjusted.Death.Rate..16.Categories..in.ranges.)
# est_death_rate is age adjusted death rate, deaths per 100,000 U.S. standard population for 2000
```

```{r}
# data cleaning
thirty_plus = which(death0$est_death_rate == "30+")
imputed_deaths = as.character(death0$est_death_rate)
for(idx in thirty_plus){
  random_death_rate = round(runif(1, 30, 957), 0) # this works now but I think we could
  # also use another (truncated) distribution that reflects distribution of adjusted death rates in
  # wv. I (revo) can look into this later if we want to.
  new_interval = paste0(random_death_rate, "-", random_death_rate)
  imputed_deaths[idx] = as.character(new_interval)
}

death0$imput_est_death_rate_num = imputed_deaths
death0$imput_est_death_rate_cat = as.factor(imputed_deaths)
death1 <- death0 %>%
  separate(imput_est_death_rate_num, c("lower_rate", "upper_rate"), sep = "-") %>%
  mutate(imput_est_death_rate_num = (as.numeric(lower_rate) + as.numeric(upper_rate))/2,
         pop_num = as.numeric(Population),
         County_name = str_replace_all(County, " County, WV", ""),
         COUNTY = str_to_upper(County_name)) %>%
  mutate(est_death_rate_cat = case_when(
    est_death_rate == "2-3.9" ~ 1,
    est_death_rate == "4-5.9" ~ 1,
    est_death_rate == "6-7.9" ~ 1,
    est_death_rate == "8-9.9" ~ 2,
    est_death_rate == "10-11.9" ~ 2,
    est_death_rate == "12-13.9" ~ 2,
    est_death_rate == "14-15.9" ~ 3,
    est_death_rate == "16-17.9" ~ 3,
    est_death_rate == "18-19.9" ~ 3,
    est_death_rate == "20-21.9" ~ 4,
    est_death_rate == "22-23.9" ~ 4,
    est_death_rate == "24-25.9" ~ 4,
    est_death_rate == "26-27.9" ~ 5,
    est_death_rate == "28-29.9" ~ 5,
    est_death_rate == "30+" ~ 6
  )) %>%
  dplyr::select(COUNTY, FIPS,Year, est_death_rate_cat, Population)
      # , est_death_rate,imput_est_death_rate_cat, imput_est_death_rate_num)
# # recreate new variables
# death2 <- death1 %>%
#   group_by(COUNTY) %>%
#   summarise(avg_death_rate = mean(est_death_rate_cat),
#             avg_pop = mean(pop_num), total_population = sum(pop_num))
```


```{r}
death1 %>%
  dplyr::select(Year, est_death_rate_cat) %>%
  mutate(Year = as.factor(Year),
         est_death_rate_cat = as.factor(est_death_rate_cat)) %>%
  ggplot(., aes(x = Year, fill = est_death_rate_cat)) +
  geom_bar(position=position_dodge())
```

<!-- ### Population data -->
<!-- ```{r} -->
<!-- pop_data = read.csv("data/2012-2018 pop for wv.csv") %>% -->
<!--   dplyr::select(GEO.display.label, respop72012) %>% -->
<!--   slice(-1) %>% -->
<!--   mutate(County = str_remove_all(GEO.display.label, "County, West Virginia"), -->
<!--          County = str_trim(County), -->
<!--          County = as.factor(str_to_upper(County)), -->
<!--          respop72012 = as.numeric(as.character(respop72012))) %>% -->
<!--   mutate(population = respop72012) %>% -->
<!--   dplyr::select(-c(GEO.display.label,respop72012)) -->
<!-- ``` -->

### Arcos dataset
```{r}
arcos0a = read.csv("arcos-wv-statewide-itemizedexcel.csv", header = TRUE)
acros0b = read_tsv("arcos-oh-statewide-itemized.tsv", col_names = TRUE)

# create subset with less variables
arcos1 <- rbind(arcos0a, acros0b) %>%
 # mutate(Year = str_sub(TRANSACTION_DATE, -4, -1)) %>%
  transmute(Year = as.numeric(str_sub(TRANSACTION_DATE, -4, -1)),
            REPORTER_NAME = as.factor(REPORTER_NAME), # name of distributor/manufacturer who reported the transaction to DEA
            REPORTER_STATE = as.factor(REPORTER_STATE),
            BUYER_TYPE = as.factor(BUYER_BUS_ACT), # types of retailers
            BUYER_NAME = as.factor(BUYER_NAME),
            BUYER_ADDRESS = as.factor(BUYER_ADDRESS1),
            BUYER_CITY = as.factor(BUYER_CITY),
            BUYER_STATE = as.factor(BUYER_STATE),
            BUYER_ZIP = as.factor(BUYER_ZIP),
            BUYER_COUNTY = as.factor(BUYER_COUNTY),
            DRUG_CODE = as.factor(DRUG_CODE),
            DRUG_NAME = as.factor(DRUG_NAME),
            QUANTITY = as.integer(QUANTITY),
            TOTAL_ACTIVE_WT = as.numeric(CALC_BASE_WT_IN_GM), # weight of total active drug in each transaction, in grams 
            DOSAGE_UNIT= as.numeric(DOSAGE_UNIT), # num of tablets in each unit
            PRODUCT_NAME = as.factor(Product_Name),
            MEASURE = as.factor(Measure), # form of drugs - in WV dataset, all are pills
            EQU_MORPHINE = as.numeric(MME_Conversion_Factor), # equivalent amt of morphine in mg
            MANUFACTURER = as.factor(Revised_Company_Name), # entity that manufactured, distributed or 
            # relabeled the drug product in the transaction
            DOS_STR = as.numeric(dos_str)) # strength of each dosage in mg (in this case, strength in each tablet)
# head(arcos1)
# calculate total active weight in each county from 2006-2012
all_act_wt <- arcos1 %>%
  group_by(BUYER_COUNTY) %>%
  summarise(all_active_wt = sum(TOTAL_ACTIVE_WT))
```

```{r}
# death rate EDA
death1 %>%
  ggplot()+
  facet_wrap(. ~ COUNTY) +
  geom_col(aes(x = Year, y = Population)) +
  labs(title = "County Population from 2006 - 2012, WV") # population is almost constant over the seven years

death1 %>%
  ggplot()+
  facet_wrap(. ~ COUNTY) +
  geom_col(aes(x = Year, y = est_death_rate_cat)) +
  # geom_hline(yintercept = median(death_rate))+
  labs(title = "Drug-related Death Rate from 2006 - 2012, WV") # death rate in general has a upward trend

# arcos original dataset EDA
# total active weight by county and drug name
arcos_wt_eda <- arcos1 %>%
  group_by(BUYER_COUNTY, Year) %>%
  summarise(active_wt_by_year = sum(TOTAL_ACTIVE_WT))

# head(arcos_wt_eda)

arcos_wt_eda %>%
  ggplot()+
  facet_wrap(. ~ BUYER_COUNTY) +
  geom_col(aes(x = Year, y = active_wt_by_year)) +
  labs(title = "Total Active Drug Weight (in grams) from 2006-2012 by County, WV") # total active wt in general has an upward trend
```

### Income dataset
```{r}
wv_income_data = read.csv("data/wv_median_household_income_data.csv") %>%
  filter(year == 2012) %>%
  filter(county != "West Virginia") %>%
  mutate(county = str_to_upper(county)) %>%
  dplyr::select(county, median_income)

oh_income_data = read.csv("data/ohio_median_household_income_data.csv") %>%
  dplyr::select(GCT_STUB.display.label.1, HC01) %>%
  clean_names() %>%
  transmute(county = str_trim(str_to_upper(str_remove_all(as.character(gct_stub_display_label_1), " County"))),
            median_income = as.numeric(as.character(hc01))) %>%
  slice(-1,-2)

income_data = rbind(wv_income_data, oh_income_data)
```

### Party affiliation
```{r}
wv_pol_affiliation = read.csv("data/wv_politics.csv")
oh_pol_affilitation = read.csv("data/oh_politics.csv") %>%
  mutate(Political.affiliation = Party.affiliation,
         State = str_trim(str_to_upper(State))) %>%
  dplyr::select(-Party.affiliation) %>%
  filter(State != "")

pol_affiliation = rbind(wv_pol_affiliation, oh_pol_affilitation)
```

### Merge datasets
```{r}
# merge arcos1 and death2
arcos_and_death <- merge(arcos1, death1, by.x = c("BUYER_COUNTY", "Year"), by.y = c("COUNTY", "Year"), all.x = TRUE)
# merge income with arcos and death data
arcos_death_and_income = merge(arcos_and_death, income_data, by.x = "BUYER_COUNTY", by.y = "county", all.x = TRUE)
# merge arcos_death_and_income with policatal affiliation data
wv_all0 = merge(arcos_death_and_income, pol_affiliation, by.x = "BUYER_COUNTY", by.y = "State", all.x = TRUE)
# merge with population data, create total active weight in county per person
# wv_all1 = merge(wv_all0, pop_data, by.x = "BUYER_COUNTY", by.y = "County", all.x = TRUE)

TAWICPP <- wv_all0 %>%
  group_by(BUYER_COUNTY) %>%
  summarise(county_total_wt=sum(TOTAL_ACTIVE_WT), Population=first(Population)) %>%
  mutate(act_wt_person_county = county_total_wt/Population)

oh_wv_all = merge(wv_all0, TAWICPP, by.x = "BUYER_COUNTY", by.y = "BUYER_COUNTY", all.x = TRUE) %>%
  filter(Year == 2012) %>% # only consider 2012
  filter(!is.na(act_wt_person_county))
```

## Exploratory Data Analysis

```{r}
wt_p = ggplot(oh_wv_all, aes(x = TOTAL_ACTIVE_WT)) +
  geom_histogram() +
  labs(x = "Total Active Weight")
twt_p = ggplot(oh_wv_all, aes(x = log(TOTAL_ACTIVE_WT))) +
  geom_histogram() +
  labs(x = "Log of Total Active Weight")
plot_grid(wt_p, twt_p)
du_p = ggplot(oh_wv_all, aes(x = DOSAGE_UNIT)) +
  geom_histogram() +
  labs(x = "Dosage Unit")
tdu_p = ggplot(oh_wv_all, aes(x = log(DOSAGE_UNIT))) +
  geom_histogram() +
  labs(x = "Log Dosage Unit")
plot_grid(du_p, tdu_p)
ds_p = ggplot(oh_wv_all, aes(x = DOS_STR)) +
  geom_histogram() +
  labs(x = "Dosage Strength")
ds_p
```

Average active weight of drugs per person by county plotted on a map and compared to opioids deaths by county.
```{r}
ggplot(data = oh_wv_all) +
  geom_point(aes(x = act_wt_person_county, y = est_death_rate_cat)) +
  labs(x = "Total Active Weight of Drug Per Person 2006-2012",
       y = "Total Overdose Death Rate in a County 2006-2012")
```

Relationship between number of manufacturers/distributors shipping to a county vs opioid deaths by county.
```{r}
oh_wv_all %>%
  group_by(REPORTER_NAME, est_death_rate_cat, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY, est_death_rate_cat) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters)) %>%
  ggplot(., aes(x = est_death_rate_cat, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Average death rate of a county", y = "Number of manufacturers involved",
       title = "Relationship between number of manufacturers and death rates in a county?")
```

Number of manufacturers/distributors shipping to a county vs amount of opioids in active weight ordered by 
pharmacies in the county.
```{r}
oh_wv_all %>%
  group_by(REPORTER_NAME, TOTAL_ACTIVE_WT, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters),
            total_active_drug_for_given_county = sum(TOTAL_ACTIVE_WT)) %>%
  ggplot(., aes(x = total_active_drug_for_given_county, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Total Active Drug Weight", y = "Number of manufacturers shipping to the county",
       title = "Total active drug weight vs. number of manufacturers for that county")
```

Average active weight of drugs per person vs number of pharmacies per 10,000 people.
```{r}
pharmacies = oh_wv_all %>%
  distinct(BUYER_ADDRESS, .keep_all = TRUE) %>%
  group_by(BUYER_COUNTY) %>%
  summarize(number_of_pharmacies = n(), Population = first(Population))

active_weight = oh_wv_all %>%
  group_by(BUYER_COUNTY) %>%
  distinct(act_wt_person_county, .keep_all = TRUE) %>%
  ungroup() %>%
  dplyr::select(BUYER_COUNTY, act_wt_person_county)
  
pop_and_pharmacies = inner_join(active_weight, pharmacies) %>%
  ###### someone check if this is how you calculate number of pharmacies per 10,000 ######
  mutate(pharmacies_per_tt = (10000*number_of_pharmacies)/Population)

  ########################################################################################
ggplot(pop_and_pharmacies, aes(x = act_wt_person_county, y = pharmacies_per_tt)) +
geom_point() +
labs(x = "Avg. active weight of drugs per person", y = "Number of pharmacies")
```

Distribution of oxycodone and hydrocodone.
```{r}
```

Average active weight of drugs per person vs strength of dose.
```{r}
# someone has code for this I think
```

```{r}
# total active weight by county and drug name
arcos2 <- arcos1 %>%
  group_by(BUYER_COUNTY,DRUG_NAME) %>%
  summarise(all_active_wt = sum(TOTAL_ACTIVE_WT)) 
head(arcos2)
```

## Modeling:
```{r}
```


