---
title: "STA325 Final Project"
author: "Alice Liao, Revo Tesha, Salvador, Cindy, Evelyn "
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(cowplot)
library(janitor)
set.seed(10)
```

## Data

### Drug poisoning dataset
```{r}
# load original dataset
death.wv <- read.csv("data/Drug_Poisoning_Mortality_by_County.csv", header = TRUE) 

death.oh <- read.csv("data/NCHS_-_Drug_Poisoning_Mortality_by_County__United_Statesoh.csv", 
                     header = TRUE, as.is = TRUE) %>%
   mutate(Population = as.numeric(gsub(",","",Population,fixed=TRUE)))

death0<-rbind(death.wv, death.oh)%>%
    rename(est_death_rate = Estimated.Age.adjusted.Death.Rate..16.Categories..in.ranges.)
# est_death_rate is age adjusted death rate, deaths per 100,000 U.S. standard population for 2000
```

```{r}
# data cleaning
thirty_plus = which(death0$est_death_rate == "30+")
imputed_deaths = as.character(death0$est_death_rate)
for(idx in thirty_plus){
  random_death_rate = round(runif(1, 30, 957), 0) # this works now but I think we could
  # also use another (truncated) distribution that reflects distribution of adjusted death rates in
  # wv. I (revo) can look into this later if we want to.
  new_interval = paste0(random_death_rate, "-", random_death_rate)
  imputed_deaths[idx] = as.character(new_interval)
}

death0$imput_est_death_rate_num = imputed_deaths
death0$imput_est_death_rate_cat = as.factor(imputed_deaths)
death1 <- death0 %>%
  separate(imput_est_death_rate_num, c("lower_rate", "upper_rate"), sep = "-") %>%
  mutate(imput_est_death_rate_num = (as.numeric(lower_rate) + as.numeric(upper_rate))/2,
         pop_num = as.numeric(Population),
         County_name = case_when(State == "West Virginia" ~ str_replace_all(County, " County, WV", ""),
                                 TRUE ~ str_replace_all(County, " County, OH", "")),
         COUNTY = str_to_upper(County_name),
         STATE = case_when(State == "West Virginia" ~ "WV",
                           TRUE ~ "OH"),
         est_death_rate_cat = case_when(
           est_death_rate %in% c("2-3.9", "4-5.9","6-7.9")  ~ 1,
           est_death_rate %in% c("8-9.9", "10-11.9","12-13.9")  ~ 2,
           est_death_rate %in% c("14-15.9", "16-17.9", "18-19.9") ~ 3,
           est_death_rate %in% c("20-21.9", "22-23.9", "24-25.9") ~ 4,
           est_death_rate %in% c("26-27.9","28-29.9") ~ 5,
           TRUE ~ 6))%>%
  select(STATE, COUNTY, FIPS,Year, est_death_rate_cat, Population)
      # , est_death_rate,imput_est_death_rate_cat, imput_est_death_rate_num)
# # recreate new variables
# death2 <- death1 %>%
#   group_by(COUNTY) %>%
#   summarise(avg_death_rate = mean(est_death_rate_cat),
#             avg_pop = mean(pop_num), total_population = sum(pop_num))
```


```{r}
death1 %>%
  dplyr::select(Year, est_death_rate_cat) %>%
  mutate(Year = as.factor(Year),
         est_death_rate_cat = as.factor(est_death_rate_cat)) %>%
  ggplot(., aes(x = Year, fill = est_death_rate_cat)) +
  geom_bar(position=position_dodge())
```

### Arcos dataset
```{r}
arcos0a = read.csv("arcos-wv-statewide-itemizedexcel.csv", header = TRUE)
acros0b = read_tsv("arcos-oh-statewide-itemized.tsv", col_names = TRUE)

# create subset with less variables
arcos1 <- rbind(arcos0a, acros0b) %>%
 # mutate(Year = str_sub(TRANSACTION_DATE, -4, -1)) %>%
  transmute(Year = as.numeric(str_sub(TRANSACTION_DATE, -4, -1)),
            REPORTER_NAME = as.factor(REPORTER_NAME), # name of distributor/manufacturer who reported the transaction to DEA
            REPORTER_STATE = as.factor(REPORTER_STATE),
            BUYER_TYPE = as.factor(BUYER_BUS_ACT), # types of retailers
            BUYER_NAME = as.factor(BUYER_NAME),
            BUYER_ADDRESS = as.factor(BUYER_ADDRESS1),
            BUYER_CITY = as.factor(BUYER_CITY),
            BUYER_STATE = as.factor(BUYER_STATE),
            BUYER_ZIP = as.factor(BUYER_ZIP),
            BUYER_COUNTY = as.factor(BUYER_COUNTY),
            DRUG_CODE = as.factor(DRUG_CODE),
            DRUG_NAME = as.factor(DRUG_NAME),
            QUANTITY = as.integer(QUANTITY),
            TOTAL_ACTIVE_WT = as.numeric(CALC_BASE_WT_IN_GM), # weight of total active drug in each transaction, in grams 
            DOSAGE_UNIT= as.numeric(DOSAGE_UNIT), # num of tablets in each unit
            PRODUCT_NAME = as.factor(Product_Name),
            MEASURE = as.factor(Measure), # form of drugs - in WV dataset, all are pills
            EQU_MORPHINE = as.numeric(MME_Conversion_Factor), # equivalent amt of morphine in mg
            MANUFACTURER = as.factor(Revised_Company_Name), # entity that manufactured, distributed or 
            # relabeled the drug product in the transaction
            DOS_STR = as.numeric(dos_str)) # strength of each dosage in mg (in this case, strength in each tablet)
head(arcos1)
```

```{r}
# death rate EDA
death1 %>%
  ggplot()+
  facet_wrap(. ~ COUNTY) +
  geom_col(aes(x = Year, y = Population)) +
  labs(title = "County Population from 2006 - 2012, WV") # population is almost constant over the seven years

death1 %>%
  ggplot()+
  facet_wrap(. ~ COUNTY) +
  geom_col(aes(x = Year, y = est_death_rate_cat)) +
  # geom_hline(yintercept = median(death_rate))+
  labs(title = "Drug-related Death Rate from 2006 - 2012, WV") # death rate in general has a upward trend

# arcos original dataset EDA
# total active weight by county and drug name
arcos_wt_eda <- arcos1 %>%
  group_by(BUYER_COUNTY, Year) %>%
  summarise(active_wt_by_year = sum(TOTAL_ACTIVE_WT))

# head(arcos_wt_eda)

arcos_wt_eda %>%
  ggplot()+
  facet_wrap(. ~ BUYER_COUNTY) +
  geom_col(aes(x = Year, y = active_wt_by_year)) +
  labs(title = "Total Active Drug Weight (in grams) from 2006-2012 by County, WV") # total active wt in general has an upward trend
```

### Income dataset
```{r}
wv_income_data = read.csv("data/wv_median_household_income_data.csv") %>%
  filter(year == 2012) %>%
  filter(county != "West Virginia") %>%
  mutate(county = str_to_upper(county),
         state = "WV") %>%
  dplyr::select(state, county, median_income)

oh_income_data = read.csv("data/ohio_median_household_income_data.csv") %>%
  dplyr::select(GCT_STUB.display.label.1, HC01) %>%
  clean_names() %>% # need library(janitor)
  transmute(state = "OH",
            county = str_trim(str_to_upper(str_remove_all(as.character(gct_stub_display_label_1), " County"))),
            median_income = as.numeric(as.character(hc01))) %>%
  slice(-1,-2)

income_data = rbind(wv_income_data, oh_income_data)
```

### Party affiliation
```{r}
wv_pol_affiliation = read.csv("data/wv_politics.csv") %>%
  transmute(state = "WV",
            county = State,
            Political.affiliation = Political.affiliation)
oh_pol_affilitation <- read.csv("data/oh_politics.csv") %>%
  transmute(state = "OH",
            county = State,
            Political.affiliation = Party.affiliation) %>%
  filter(county != "")
pol_affiliation <- rbind(wv_pol_affiliation, oh_pol_affilitation)
  
```

### Merge datasets
```{r}
# arcos collapse by county, create relevant variables, only consider 2012
arcos2012 <- arcos1 %>%
  filter(Year == "2012")

# calculate total active weight in each county 2012
all_act_wt <- arcos2012 %>%
  group_by(BUYER_STATE, BUYER_COUNTY) %>%
  summarise(all_active_wt = sum(TOTAL_ACTIVE_WT))

# New Variable: % oxycodone and % hydrocodone in each count
oxy_sum <- arcos2012%>%
  filter(DRUG_NAME == "OXYCODONE") %>%
  group_by(BUYER_STATE,BUYER_COUNTY) %>%
  summarise(oxy_wt = sum(TOTAL_ACTIVE_WT))
head(oxy_sum)

hyd_sum <- arcos2012%>%
  filter(DRUG_NAME == "HYDROCODONE") %>%
  group_by(BUYER_STATE,BUYER_COUNTY) %>%
  summarise(hyd_wt = sum(TOTAL_ACTIVE_WT))
head(hyd_sum)

perc_drug <- merge(oxy_sum, hyd_sum, by.x = c("BUYER_STATE","BUYER_COUNTY"), by.y = c("BUYER_STATE","BUYER_COUNTY"), all.x = TRUE) %>%
  mutate(perc_oxy = round(oxy_wt/(oxy_wt+hyd_wt),3)*100,
         perc_hyd = 100-perc_oxy)

# New Variable: most common distribution channel (by number of transactions) in each county
most_common_dist <- arcos2012 %>%
  group_by(BUYER_STATE,BUYER_COUNTY) %>%
  summarise(perc_retail = sum(BUYER_TYPE== "RETAIL PHARMACY")/n(), 
            perc_chain = sum(BUYER_TYPE == "CHAIN PHARMACY")/n(),
            perc_practitioner = sum(BUYER_TYPE %in% c("PRACTITIONER", "PRACTITIONER-DW/100",
                                                     "PRACTITIONER-DW/275","PRACTITIONER-DW/30"))/n())%>%
  mutate(most_dist_channel = case_when(
    perc_retail >= perc_chain & perc_retail >= perc_practitioner ~ "RETAIL PHARMACY",
    perc_chain >=  perc_retail & perc_chain >= perc_practitioner ~ "CHAIN PHARMACY",
    TRUE ~ "PRACTITIONER")) 
head(most_common_dist)

# New Variable: number of pharmacies in each county that buy drugs from manufacturers
pharm_num <- arcos2012 %>%
  group_by(BUYER_STATE,BUYER_COUNTY) %>%
  summarise(pharmacy_num = n_distinct(BUYER_ADDRESS))
head(pharm_num)

# New Variable: market dominance, "yes" - if two pharmarcies in a county account for more than 50% of total active weight of drugs
market_dom <- arcos2012 %>%
  group_by(BUYER_STATE,BUYER_COUNTY, BUYER_ADDRESS) %>%
  summarise(pharm_wt = sum(TOTAL_ACTIVE_WT)) %>%
  mutate(market_share = round(pharm_wt/sum(pharm_wt),3)*100) %>%
  arrange(BUYER_COUNTY, desc(market_share)) %>%
  mutate(dominance = "No")

for(county in unique(market_dom$BUYER_COUNTY)){
  index <- which(market_dom$BUYER_COUNTY == county)
  c.df <- market_dom %>%
    filter(BUYER_COUNTY == county)
  if(sum(c.df$market_share[1:2]) >50 & length(c.df$market_share) >=2){
    market_dom$dominance[index] <- "Yes"}
  else if(length(c.df$market_share)==1){
    market_dom$dominance[index] <- "Yes"}
}
market_dom2 <- market_dom %>%
  select(BUYER_STATE,BUYER_COUNTY, dominance) %>%
  distinct()
```
```{r}

```

#### EACH ROW OF oh_wv_all IS STILL A TRANSACTION 

#### there is problem with this merger - two counties in different states might have the same name
```{r eval=FALSE}
# merge arcos1 and death1
arcos_and_death <- merge(arcos1, death1, by.x = c("BUYER_COUNTY", "Year"), by.y = c("COUNTY", "Year"), all.x = TRUE)
# merge income with arcos and death data
arcos_death_and_income = merge(arcos_and_death, income_data, by.x = "BUYER_COUNTY", by.y = "county", all.x = TRUE)
# merge arcos_death_and_income with policatal affiliation data
all0 = merge(arcos_death_and_income, pol_affiliation, by.x = "BUYER_COUNTY", by.y = "State", all.x = TRUE)
# merge with population data, create total active weight in county per person

TAWICPP <- all0 %>%
  group_by(BUYER_COUNTY) %>%
  summarise(county_total_wt=sum(TOTAL_ACTIVE_WT), Population=first(Population)) %>%
  mutate(act_wt_person_county = county_total_wt/Population)

oh_wv_all = merge(all0, TAWICPP, by.x = "BUYER_COUNTY", by.y = "BUYER_COUNTY", all.x = TRUE) %>%
  filter(Year == 2012) %>% # only consider 2012
  filter(!is.na(act_wt_person_county))
```

#### EACH ROW IS A COUNTY
```{r}
# recreate new arcos dataset - 2012 data with relevant variables
arcos2.1 <- merge(all_act_wt, perc_drug, by.x = c("BUYER_STATE","BUYER_COUNTY"), by.y = c("BUYER_STATE", "BUYER_COUNTY"), all.x = TRUE)
arcos2.2 <- merge(arcos2.1, most_common_dist, by.x = c("BUYER_STATE","BUYER_COUNTY"), by.y = c("BUYER_STATE", "BUYER_COUNTY"), all.x = TRUE)
arcos2.3 <- merge(arcos2.2, pharm_num, by.x = c("BUYER_STATE","BUYER_COUNTY"), by.y = c("BUYER_STATE", "BUYER_COUNTY"), all.x = TRUE)
arcos_county <- merge(arcos2.3, market_dom2, by.x = c("BUYER_STATE","BUYER_COUNTY"), by.y = c("BUYER_STATE", "BUYER_COUNTY"), all.x = TRUE)

death2012 <- death1 %>%
  filter(Year == "2012")
  
# merge arcos_county and death2012
arcos_and_death2012 <- merge(arcos_county, death2012, by.x = c("BUYER_STATE","BUYER_COUNTY"), by.y = c("STATE", "COUNTY"), all.x = TRUE) %>%
  select(-FIPS)

# merge income with arcos_county and death2012 and income data
arcos_death_and_income2012 = merge(arcos_and_death2012, income_data, 
                                   by.x = c("BUYER_STATE", "BUYER_COUNTY"), by.y = c("state", "county"), all.x = TRUE)

# merge arcos_death_and_income2012 with policatal affiliation data
all2012 = merge(arcos_death_and_income2012, pol_affiliation, 
                by.x = c("BUYER_STATE", "BUYER_COUNTY"), by.y = c("state","county"), all.x = TRUE)

# merge with population data, create total active weight in county per person
oh_wv_2012 <- all2012 %>%
  mutate(act_wt_person_county = all_active_wt/Population)

```

```{r}
# export oh_wv_all and oh_wv_2012 to csv file
#write_csv(oh_wv_all, "oh_wv_all.csv")
write_csv(oh_wv_2012, "oh_wv_2012.csv")
```

## Exploratory Data Analysis

```{r}
wt_p = ggplot(oh_wv_all, aes(x = TOTAL_ACTIVE_WT)) +
  geom_histogram() +
  labs(x = "Total Active Weight")
twt_p = ggplot(oh_wv_all, aes(x = log(TOTAL_ACTIVE_WT))) +
  geom_histogram() +
  labs(x = "Log of Total Active Weight")
plot_grid(wt_p, twt_p)
du_p = ggplot(oh_wv_all, aes(x = DOSAGE_UNIT)) +
  geom_histogram() +
  labs(x = "Dosage Unit")
tdu_p = ggplot(oh_wv_all, aes(x = log(DOSAGE_UNIT))) +
  geom_histogram() +
  labs(x = "Log Dosage Unit")
plot_grid(du_p, tdu_p)
ds_p = ggplot(oh_wv_all, aes(x = DOS_STR)) +
  geom_histogram() +
  labs(x = "Dosage Strength")
ds_p
```

Average active weight of drugs per person by county plotted on a map and compared to opioids deaths by county.
```{r}
ggplot(data = oh_wv_all) +
  geom_point(aes(x = act_wt_person_county, y = est_death_rate_cat)) +
  labs(x = "Total Active Weight of Drug Per Person 2006-2012",
       y = "Total Overdose Death Rate in a County 2006-2012")
```

Relationship between number of manufacturers/distributors shipping to a county vs opioid deaths by county.
```{r}
oh_wv_all %>%
  group_by(REPORTER_NAME, est_death_rate_cat, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY, est_death_rate_cat) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters)) %>%
  ggplot(., aes(x = est_death_rate_cat, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Average death rate of a county", y = "Number of manufacturers involved",
       title = "Relationship between number of manufacturers and death rates in a county?")
```

Number of manufacturers/distributors shipping to a county vs amount of opioids in active weight ordered by 
pharmacies in the county.
```{r}
oh_wv_all %>%
  group_by(REPORTER_NAME, TOTAL_ACTIVE_WT, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters),
            total_active_drug_for_given_county = sum(TOTAL_ACTIVE_WT)) %>%
  ggplot(., aes(x = total_active_drug_for_given_county, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Total Active Drug Weight", y = "Number of manufacturers shipping to the county",
       title = "Total active drug weight vs. number of manufacturers for that county")
```

Average active weight of drugs per person vs number of pharmacies per 10,000 people.
```{r}
pharmacies = oh_wv_all %>%
  distinct(BUYER_ADDRESS, .keep_all = TRUE) %>%
  group_by(BUYER_COUNTY) %>%
  summarize(number_of_pharmacies = n(), Population = first(Population))

active_weight = oh_wv_all %>%
  group_by(BUYER_COUNTY) %>%
  distinct(act_wt_person_county, .keep_all = TRUE) %>%
  ungroup() %>%
  dplyr::select(BUYER_COUNTY, act_wt_person_county)
  
pop_and_pharmacies = inner_join(active_weight, pharmacies) %>%
  ###### someone check if this is how you calculate number of pharmacies per 10,000 ######
  mutate(pharmacies_per_tt = (10000*number_of_pharmacies)/Population)

  ########################################################################################
ggplot(pop_and_pharmacies, aes(x = act_wt_person_county, y = pharmacies_per_tt)) +
geom_point() +
labs(x = "Avg. active weight of drugs per person", y = "Number of pharmacies")
```

Distribution of oxycodone and hydrocodone.
```{r}
```

Average active weight of drugs per person vs strength of dose.
```{r}
# someone has code for this I think
```

```{r}
# total active weight by county and drug name
arcos2 <- arcos1 %>%
  group_by(BUYER_COUNTY,DRUG_NAME) %>%
  summarise(all_active_wt = sum(TOTAL_ACTIVE_WT)) 
head(arcos2)
```



