---
title: "STA325 Final Project"
author: "Alice Liao, Revo Tesha, Salvador ... "
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
set.seed(10)
```

## Data

### Drug poisoning dataset
```{r}
# load original dataset
death0 <- read.csv("data/Drug_Poisoning_Mortality_by_County.csv", header = TRUE) %>%
    rename(est_death_rate = Estimated.Age.adjusted.Death.Rate..16.Categories..in.ranges.)
# est_death_rate is age adjusted death rate, deaths per 100,000 U.S. standard population for 2000
```

```{r}
# data cleaning
thirty_plus = which(death0$est_death_rate == "30+")
imputed_deaths = as.character(death0$est_death_rate)

for(idx in thirty_plus){
  random_death_rate = round(runif(1, 30, 957), 0) # this works now but I think we could
  # also use another (truncated) distribution that reflects distribution of adjusted death rates in
  # wv. I (revo) can look into this later if we want to.
  new_interval = paste0(random_death_rate, "-", random_death_rate)
  imputed_deaths[idx] = as.character(new_interval)
}

death0$imput_est_death_rate_num = imputed_deaths
death0$imput_est_death_rate_cat = as.factor(imputed_deaths)

death1 <- death0 %>%
  separate(imput_est_death_rate_num, c("lower_rate", "upper_rate"), sep = "-") %>%
  mutate(imput_est_death_rate_num = (as.numeric(lower_rate) + as.numeric(upper_rate))/2,
         pop_num = as.numeric(Population),
         County_name = str_replace_all(County, " County, WV", ""),
         COUNTY = str_to_upper(County_name)) %>%
  mutate(est_death_rate_cat = case_when(
    est_death_rate == "2-3.9" ~ 1,
    est_death_rate == "4-5.9" ~ 1,
    est_death_rate == "6-7.9" ~ 1,
    est_death_rate == "8-9.9" ~ 2,
    est_death_rate == "10-11.9" ~ 2,
    est_death_rate == "12-13.9" ~ 2,
    est_death_rate == "14-15.9" ~ 3,
    est_death_rate == "16-17.9" ~ 3,
    est_death_rate == "18-19.9" ~ 3,
    est_death_rate == "20-21.9" ~ 4,
    est_death_rate == "22-23.9" ~ 4,
    est_death_rate == "24-25.9" ~ 4,
    est_death_rate == "26-27.9" ~ 5,
    est_death_rate == "28-29.9" ~ 5,
    est_death_rate == "30+" ~ 6
  )) %>%
  dplyr::select(COUNTY, FIPS,Year, est_death_rate_cat)
      # , est_death_rate,imput_est_death_rate_cat, imput_est_death_rate_num)

# # recreate new variables
# death2 <- death1 %>%
#   group_by(COUNTY) %>%
#   summarise(avg_death_rate = mean(est_death_rate_cat),
#             avg_pop = mean(pop_num), total_population = sum(pop_num))
```


```{r}
# does distribution of estimated death rate change by year? looks like it doesnt.
# so we pick one year? continuing with all years included for now.
death1 %>%
  dplyr::select(Year, est_death_rate_cat) %>%
  mutate(Year = as.factor(Year),
         est_death_rate_cat = as.factor(est_death_rate_cat)) %>%
  ggplot(., aes(x = Year, fill = est_death_rate_cat)) +
  geom_bar(position=position_dodge())
```

### Population data
```{r}
pop_data = read.csv("data/2012-2018 pop for wv.csv") %>%
  dplyr::select(GEO.display.label, respop72010, respop72011, respop72012) %>%
  slice(-1) %>%
  mutate(County = str_remove_all(GEO.display.label, "County, West Virginia"),
         County = str_trim(County),
         County = as.factor(str_to_upper(County)),
         respop72010 = as.numeric(as.character(respop72010)),
         respop72011 = as.numeric(as.character(respop72011)),
         respop72012 = as.numeric(as.character(respop72012))) %>%
  mutate(population = (respop72010 + respop72011 + respop72012)/3) %>%
  dplyr::select(-c(GEO.display.label, respop72010, respop72011, respop72012))
```

### Arcos dataset
```{r}
arcos0 <- read.csv("arcos-wv-statewide-itemizedexcel.csv", header = TRUE)

# create subset with less variables
arcos1 <- arcos0 %>%
 # mutate(Year = str_sub(TRANSACTION_DATE, -4, -1)) %>%
  transmute(Year = as.numeric(str_sub(TRANSACTION_DATE, -4, -1)),
            REPORTER_NAME = as.factor(REPORTER_NAME), # name of distributor/manufacturer who reported the transaction to DEA
            REPORTER_STATE = as.factor(REPORTER_STATE),
            BUYER_TYPE = as.factor(BUYER_BUS_ACT), # types of retailers
            BUYER_NAME = as.factor(BUYER_NAME),
            BUYER_ADDRESS = as.factor(BUYER_ADDRESS1),
            BUYER_CITY = as.factor(BUYER_CITY),
            BUYER_STATE = as.factor(BUYER_STATE),
            BUYER_ZIP = as.factor(BUYER_ZIP),
            BUYER_COUNTY = as.factor(BUYER_COUNTY),
            DRUG_CODE = as.factor(DRUG_CODE),
            DRUG_NAME = as.factor(DRUG_NAME),
            QUANTITY = as.integer(QUANTITY),
            TOTAL_ACTIVE_WT = as.numeric(CALC_BASE_WT_IN_GM), # weight of total active drug in each transaction, in grams 
            DOSAGE_UNIT= as.numeric(DOSAGE_UNIT), # num of tablets in each unit
            PRODUCT_NAME = as.factor(Product_Name),
            MEASURE = as.factor(Measure), # form of drugs - in WV dataset, all are pills
            EQU_MORPHINE = as.numeric(MME_Conversion_Factor), # equivalent amt of morphine in mg
            MANUFACTURER = as.factor(Revised_Company_Name), # entity that manufactured, distributed or 
            # relabeled the drug product in the transaction
            DOS_STR = as.numeric(dos_str)) # strength of each dosage in mg (in this case, strength in each tablet)
# head(arcos1)
# calculate total active weight in each county from 2006-2012
all_act_wt <- arcos1 %>%
  group_by(BUYER_COUNTY) %>%
  summarise(all_active_wt = sum(TOTAL_ACTIVE_WT))
```

### Income dataset
```{r}
# income data -- is this adjusted for inflation? also only goes from 2011 to 2016.
wv_income_data = read.csv("data/wv_median_household_income_data.csv") %>%
  filter(county != "West Virginia") %>%
  mutate(county = str_to_upper(county)) %>%
  group_by(county) %>%
  summarize(average_median_income = mean(median_income))
```

### Party affiliation
```{r}
incomp_pol_affiliation = read.csv("data/wv_politics.csv")
# according to https://www.politico.com/election-results/2018/west-virginia/,
# taylor county is republican; taylor was missing in the csv above
taylor_cnty = tibble(State = "TAYLOR", Political.affiliation = "Republican")
pol_affiliation = rbind(incomp_pol_affiliation, taylor_cnty)
```

### Merge datasets
```{r}
# merge arcos1 and death2
arcos_and_death <- merge(arcos1, death1, by.x = c("BUYER_COUNTY", "Year"), by.y = c("COUNTY", "Year"), all.x = TRUE)

# merge income with arcos and death data
arcos_death_and_income = merge(arcos_and_death, wv_income_data, by.x = "BUYER_COUNTY", by.y = "county", all.x = TRUE)

# merge arcos_death_and_income with policatal affiliation data
wv_all0 = merge(arcos_death_and_income, pol_affiliation, by.x = "BUYER_COUNTY", by.y = "State", all.x = TRUE)

# merge with population data
wv_all = merge(wv_all0, pop_data, by.x = "BUYER_COUNTY", by.y = "County", all.x = TRUE)
```

## Exploratory Data Analysis

```{r}
# summary(death0)
# summary(death)
# summary(death2)
# head(arcos0)
# head(arcos1)
# summary(arcos1)
View(wv_all)
```

```{r}
wt_p = ggplot(wv_all, aes(x = TOTAL_ACTIVE_WT)) +
  geom_histogram() +
  labs(x = "Total Active Weight")
twt_p = ggplot(wv_all, aes(x = log(TOTAL_ACTIVE_WT))) +
  geom_histogram() +
  labs(x = "Log of Total Active Weight")
plot_grid(wt_p, twt_p)

du_p = ggplot(wv_all, aes(x = DOSAGE_UNIT)) +
  geom_histogram() +
  labs(x = "Dosage Unit")

tdu_p = ggplot(wv_all, aes(x = log(DOSAGE_UNIT))) +
  geom_histogram() +
  labs(x = "Log Dosage Unit")
plot_grid(du_p, tdu_p)

ds_p = ggplot(wv_all, aes(x = DOS_STR)) +
  geom_histogram() +
  labs(x = "Dosage Strength")

ds_p
```

Average active weight of drugs per person by county plotted on a map and compared to opioids deaths by county.
```{r}
# merge all_act_wt with death_avg
all_act_wt_death <- merge(all_act_wt, death2, by.x = "BUYER_COUNTY", by.y = "COUNTY") %>%
  mutate(act_wt_person = all_active_wt/avg_pop)

ggplot(data = all_act_wt_death) +
  geom_point(aes(x = act_wt_person, y = avg_death_rate)) +
  labs(x = "Total Active Weight of Drug Per Person 2006-2012",
       y = "Total Overdose Death Rate in a County 2006-2012")
```

Relationship between number of manufacturers/distributors shipping to a county vs opioid deaths by county.
```{r}
wv_all %>%
  group_by(REPORTER_NAME, avg_death_rate, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY, avg_death_rate) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters)) %>%
  ggplot(., aes(x = avg_death_rate, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Average death rate of a county", y = "Number of manufacturers involved",
       title = "Relationship between number of manufacturers and death rates in a county?")
```

Number of manufacturers/distributors shipping to a county vs amount of opioids in active weight ordered by 
pharmacies in the county.
```{r}
wv_all %>%
  group_by(REPORTER_NAME, TOTAL_ACTIVE_WT, BUYER_COUNTY) %>%
  summarise(number_of_reporters = n()) %>%
  group_by(BUYER_COUNTY) %>%
  summarise(number_of_reporters_for_given_county = sum(number_of_reporters),
            total_active_drug_for_given_county = sum(TOTAL_ACTIVE_WT)) %>%
  ggplot(., aes(x = total_active_drug_for_given_county, y = number_of_reporters_for_given_county)) +
  geom_point() +
  labs(x = "Total Active Drug Weight", y = "Number of manufacturers shipping to the county",
       title = "Total active drug weight vs. number of manufacturers for that county")
```

Average active weight of drugs per person vs number of pharmacies per 10,000 people.
```{r}
weight_and_pop = wv_all %>%   
  group_by(BUYER_COUNTY) %>%
  summarise(total_population = first(total_population), avg_pop = first(avg_pop), 
            all_active_wt = sum(TOTAL_ACTIVE_WT)) %>%
  mutate(act_wt_person = all_active_wt/avg_pop) %>%
  dplyr::select(-c(avg_pop, all_active_wt))

pharmacies = wv_all %>%
  filter(BUYER_TYPE == "RETAIL PHARMACY" | BUYER_TYPE == "CHAIN PHARMACY") %>%
  distinct(BUYER_ADDRESS, .keep_all = TRUE) %>%
  group_by(BUYER_COUNTY) %>%
  summarize(number_of_pharmacies = n())
  
weight_pop_and_pharmacies = inner_join(weight_and_pop, pharmacies) %>%
  ###### someone check if this is how you calculate number of pharmacies per 10,000 ######
  mutate(pharmacies_per_tt = (10000*number_of_pharmacies)/total_population)
  ########################################################################################
ggplot(weight_pop_and_pharmacies, aes(x = act_wt_person, y = pharmacies_per_tt)) +
geom_point() +
labs(x = "Avg. active weight of drugs per person", y = "Number of pharmacies")
```

Distribution of oxycodone and hydrocodone.
```{r}

```

Average active weight of drugs per person vs strength of dose.
```{r}
# someone has code for this I think
```

```{r}
# total active weight by county and drug name
arcos2 <- arcos1 %>%
  group_by(BUYER_COUNTY,DRUG_NAME) %>%
  summarise(all_active_wt = sum(TOTAL_ACTIVE_WT)) 
head(arcos2)
```

## Modeling:
```{r}

```

